---
title: "Ch06FactorsandTables"
author: "Yang Ge"
date: "November 16, 2015"
output:
  html_document:
    toc: true
---

```{r setoptions, echo=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE, cache = TRUE)
library(dplyr)
```

# 6.1 factors and levels

```{r 6.1}
x <- c(5, 12, 13, 12)
xf <- factor(x)
xf
str(xf)
unclass(xf)
length(xf)
```

The factor only stores the levels not the actual data.
The levels themselves are characters like "5" rather than 5.

We can reserve a level for future use.
```{r}
x <- c(5, 12, 13, 12)
xff <- factor(x, levels = c(5, 12, 13, 88))
xff
str(xff)
unclass(xff)
xff[2] <- 88
xff
unclass(xff)
```

You cannot sneak in an "illegal" value like this.
```{r}
xff[2] <- 28
```

If we want to make this work, we have to do this:

```{r}
x <- c(5, 12, 13, 12)
xff <- factor(x, levels = c(5, 12, 13, 88, 28))
xff[2] <- 28
xff
str(xff)
```

# 6.2 Common Functions Used with Factors

## `tapply`

First is `tapply(x, f, g)`, it can return an array or a matrix.

```{r 6.2}
ages <- c(25, 26, 55, 37, 21, 42)
affils <- c("R", "D", "D", "R", "U", "D")
tapply(ages, affils, mean)
class(tapply(ages, affils, mean))
```

If we have 2 factors in `tapply`, we will get a matrix.
```{r}
d <- data.frame(list(gender=c("M","M","F","M","F","F"),
age=c(47,59,21,32,33,24),income=c(55000,88000,32450,76500,123000,45650)))
d
d$over25 <- ifelse(d$age > 25, 1, 0)
d
tapply(d$income, list(d$gender, d$over25), mean)
class(tapply(d$income, list(d$gender, d$over25), mean))
```

## `split`

returns a list.
```{r}
split(d$income, list(d$gender, d$over25))
g <- c("M","F","F","I","M","M","F")
g.split <- split(1:length(g), g)
str(g.split)
unclass(g.split)
```

## `by()`

Function `by` is an object-oriented wrapper for tapply applied to data frames.

```{r}
aba <- read.csv("abalone.data", header = FALSE)
names(aba) <- c("Gender", "Length", "Diameter", "Height", "Whole weight",
                "Shucked weight", "Viscera weight", "Shell weight", "Rings")
head(aba)
by(data = aba, aba$Gender, FUN = function(m) lm(m[,2]~m[,3]))
class(by(data = aba, aba$Gender, FUN = function(m) lm(m[,2]~m[,3])))
```

# 6.3 Working with tables

The `table` function takes one or more object which can be interprated as factor,
and return a contigency table.

```{r}
u <- c(22,8,33,6,8,29,-2)
fl <- list(c(5,12,13,12,13,5,13),c("a","bc","a","a","bc","a","a"))
tapply(u, fl, length)
table(fl)
class(table(fl))
ct <- read.table("ct.dat", header = TRUE)
ct
cttab <- table(ct)
cttab
table(c(5,12,13,12,8,5))
```

# 6.4 Other factor- and table-related functions

## `aggregate()`

`aggregate` use `tapply` on every variable of a dataframe.
Note that the 2nd `by` variable has to be a list.

```{r aggregate}
aba <- read.csv("abalone.data", header = FALSE)
names(aba) <- c("Gender", "Length", "Diameter", "Height", "Whole weight",
                "Shucked weight", "Viscera weight", "Shell weight", "Rings")
aggregate(x = aba[,-1], by = list(aba$Gender), median)
```

## `cut()`

give a data vector x and a set of bins defined by a vector b and generate a 

```{r cut}
z <- c(0.88114802, 0.28532689, 0.58647376, 0.42851862, 0.46881514,
       0.24226859, 0.05289197, 0.88035617)
binmarks <- seq(from=0.0, to=1.0, by=0.1)
g <- cut(z, binmarks, labels = F)
class(g)
```

