---
title: "Ch06FactorsandTables"
author: "Yang Ge"
date: "November 16, 2015"
output:
  html_document:
    toc: true
---

```{r setoptions, echo=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE, cache = TRUE)
library(dplyr)
```

# 6.1 factors and levels

```{r 6.1}
x <- c(5, 12, 13, 12)
xf <- factor(x)
xf
str(xf)
unclass(xf)
length(xf)
```

The factor only stores the levels not the actual data.
The levels themselves are characters like "5" rather than 5.

We can reserve a level for future use.
```{r}
x <- c(5, 12, 13, 12)
xff <- factor(x, levels = c(5, 12, 13, 88))
xff
str(xff)
unclass(xff)
xff[2] <- 88
xff
unclass(xff)
```

You cannot sneak in an "illegal" value like this.
```{r}
xff[2] <- 28
```

If we want to make this work, we have to do this:

```{r}
x <- c(5, 12, 13, 12)
xff <- factor(x, levels = c(5, 12, 13, 88, 28))
xff[2] <- 28
xff
str(xff)
```

# 6.2 Common Functions Used with Factors

## `tapply`

First is `tapply(x, f, g)`

```{r 6.2}
ages <- c(25, 26, 55, 37, 21, 42)
affils <- c("R", "D", "D", "R", "U", "D")
tapply(ages, affils, mean)
```

If we have 2 factors in `tapply`, we will get a matrix.
```{r}
d <- data.frame(list(gender=c("M","M","F","M","F","F"),
age=c(47,59,21,32,33,24),income=c(55000,88000,32450,76500,123000,45650)))
d
d$over25 <- ifelse(d$age > 25, 1, 0)
d
tapply(d$income, list(d$gender, d$over25), mean)
```

