---
title: "The Impact of Storm on Population Health and Economic Consequences"
date: "September 24, 2015"
output: 
  html_document:
    keep_md: true
---

# Synopsis

In this report, we investigate the impace of different types of weather events
on populatoin health and economic consequences.
We find "TORNADO" has the biggest impact on pupolation health, while "FLOOD"
damages the property most.

## 2 Data Processing

We will need the `knitr` to generate the report and `dplyr` to facilitate the
data processing.

```{r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r, echo=FALSE}
if (! require(dplyr)) {
    install.packages("dplyr")
}
library(dplyr)
```

We first read in the data by calling `read.csv()` function.
Please note that I am running Rstudio on a VMware virtual machine with Ubuntu
Linux. If I load the dataset in full through `knitr`, it would get crashed.
So I limit the number of row to 300000.
On the other hand, if I just load it from console, the whole dataset can be loaded.

```{r readData}
stormData <- read.csv("repdatadataStormData.csv.bz2", header = TRUE, nrows = 300000)
```

### 2.1 Data processing for population health

We use the summation of "FATALITIES" and "INJURIES" as the measure for the impact
on populatoin health. The `group_by` function inside `dplyr` are used to group
observation based on "EVTYPE". Then we make summation with each group.
Finally, we only choose the top 10 events which have
biggest impact on populatio health by using `top_n` and `order` function. And only
use them to present results.

```{r population.health}

health <- stormData %>%
  group_by(EVTYPE) %>%
  summarise(not.health = sum(FATALITIES + INJURIES)) %>%
    top_n(10, not.health)

health <- health[with(health, order(-not.health)), ]
```

### 2.2 Data processing for economic consequences

Processing economic consequences is a little bit more tricky than population health.
We found there are some empty string `""` in `stormData$PROPDMGEXP` and
`stormData$CROPDMGEXP`. So our first step is to replace those empty strings with
a question mark `"?"`.

```{r property.damage}

## replace the observation with "" in EXP to make future processing easier
levels(stormData$PROPDMGEXP)[levels(stormData$PROPDMGEXP) == ""] <- "?"
levels(stormData$CROPDMGEXP)[levels(stormData$CROPDMGEXP) == ""] <- "?"
```

Next we build a named vector `multiplier` to map exponent character to numeric
value based on the course forum discussion
[PROPDMGEXP, CROPDMGEXP, PROPDMG and CROPDMG](https://class.coursera.org/repdata-032/forum/thread?thread_id=102).

```{r}
## Make a mapping from EXP character to numerical numbers
c1 <- c("-" = 1, "?" = 1, "+" = 1)
c2 <- 10^as.numeric(0:9)
num.char <- as.character(0:9)
names(c2) <- num.char
c3 <- c('h'=100, 'H'=100, 'k'=1000, 'K'=1000, 'm'=10^6, 'M'=10^6, 'b'=10^9, 'B'=10^9)
multiplier <- c(c1, c2, c3)
```

Here is our mapping:

```{r, echo=FALSE}
multiplier
```

Then we use `PROPDMG * PROPDMGEXP` to represent property damage
and `CROPDMG * CROPDMGEXP` to represent crop damage. We use the
summation of the two variables to represent economy consequences.
Similar to population health, we first group them based on different
`EVTYPE` and make a summation within each group and
only choose the top 10 events for results
analysis and presentation.

```{r}
## Process the data
test <- stormData %>%
    group_by(EVTYPE) %>%
    select(EVTYPE, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP) %>%
    mutate(prop.damage = multiplier[PROPDMGEXP] * PROPDMG,
           crop.damage = multiplier[CROPDMGEXP] * CROPDMG,
           econ.damage = prop.damage + crop.damage) %>%
    summarise(total.prop.damage = sum(prop.damage),
              total.crop.damage = sum(crop.damage),
              total.econ.damage = sum(econ.damage)) %>%
    top_n(10, total.econ.damage)

# Sort based on damage
test <- test[with(test, order(-total.econ.damage)), ]
total.econ.damage <- test$total.econ.damage
```

## 3. Results

### 3.1 impact on population health

From the figure and table below, we can clearly see
the worst event for population health is "TORNADO".

```{r show.results.health}
barplot(height = health$not.health,
        col = "green",
        names.arg = health$EVTYPE,
        las = 3,
        main = "Total people affected by type of events",
        ylab = "No. of people")
```

```{r results='asis'}
library(xtable)
xt <- xtable(health)
print(xt, type = "html")
```


### 3.2 impact on economic consequences

From the figure and table below, we can clearly see
the worst event for economic consequences is "TORNADO".

```{r show.results.econ}
barplot(height = total.econ.damage, col = "red", names.arg = test$EVTYPE, las = 3,
        main = "Total economy damage in different types of events",
        ylab = "US dollers")
```

```{r results='asis'}
xt <- xtable(test)
print(xt, type = "html")
```
