---
title: "Ch07ExploratoryDataAnalysis"
author: "Yang Ge"
date: "10/5/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 7.5 Covariations

### 7.5.2 Two categorical variables

```{r}
ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))
```
```{r}
diamonds %>%
  count(color, cut)

diamonds %>%
  count(color, cut) %>%
  ggplot(mapping = aes(x = color, y = cut)) +
  geom_tile(mapping = aes(fill = n))

```

#### 7.5.2.1 Exercises

1. How could you rescale the count dataset above to more clearly show the
    distribution of cut within colour, or colour within cut?

* Just needs to use create a new varible, `n/sum(n)`

```{r}
diamonds %>%
  count(color, cut) %>%
  mutate(density = n / sum(n), sum = sum(n))
```

2. Use geom_tile() together with dplyr to explore how average flight delays 
   vary by destination and month of year. What makes the plot difficult to read?
   How could you improve it?
   
* I think one reason for hard to read is there are so many distinations.
* Question, I don't have good answers for the 2nd part. Maybe we can divide the destination
  according to their first letter?

```{r}
library(nycflights13)
flights
flights %>%
  group_by(dest, month) %>%
  summarise(mean = mean(arr_delay, na.rm = TRUE)) %>%
  filter(dest < "B")

flights %>%
  group_by(dest, month) %>%
  summarise(mean = mean(arr_delay, na.rm = TRUE)) %>%
  ggplot(mapping = aes(x = month, y = dest)) +
  geom_tile(mapping = aes(fill = mean))
```

3. Why is it slightly better to use aes(x = color, y = cut) rather than
    aes(x = cut, y = color) in the example above?
    
* aes(x = color, y = cut) looks pretty than aes(x = cut, y = color) ...

```{r}
diamonds %>%
  count(color, cut) %>%
  ggplot(mapping = aes(x = cut, y = color)) +
  geom_tile(mapping = aes(fill = n))
```

### 7.5.3 Two continuous variables

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))
```

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price), alpha = 0.1)
```

* use `geom_bin2d`, `geom_hex`

```{r}
smaller <- diamonds %>%
  filter(carat < 3)
ggplot(data = smaller) +
  geom_bin2d(mapping = aes(x = carat, y = price))

# install.packages("hexbin")
ggplot(data = smaller) +
  geom_hex(mapping = aes(x = carat, y = price))

```

* bin one variable

```{r}
ggplot(data = smaller, mapping = aes(x = carat, y = price)) +
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)))

ggplot(data = smaller, mapping = aes(x = carat, y = price)) +
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)), varwidth = TRUE)

ggplot(data = smaller, mapping = aes(x = carat, y = price)) +
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```

#### 7.5.3.1 Exercises

1. Instead of summarising the conditional distribution with a boxplot,
  you could use a frequency polygon. What do you need to consider when
  using cut_width() vs cut_number()? How does that impact a visualisation
  of the 2d distribution of carat and price?

* For `cut_width()`, it needs to choose correct `width, center` to get reasonable # of plots.
* For `cut_number`, it needs to choose good `n`.

```{r}
ggplot(data = smaller, mapping = aes(x = price, y = ..density..)) +
  geom_freqpoly(mapping = aes(color = cut_width(carat, 1, center = 0.5)), binwidth = 500)

ggplot(data = smaller, mapping = aes(x = price, y = ..density..)) +
  geom_freqpoly(mapping = aes(color = cut_number(carat, 5)), binwidth = 500)

```

* Visualise the distribution of carat, partitioned by price.

```{r}
ggplot(data = diamonds, mapping = aes(x = price, y = carat)) +
  geom_boxplot(mapping = aes(group = cut_width(price, 1000)))
```

* How does the price distribution of very large diamonds compare to small diamonds.
  Is it as you expect, or does it surprise you?
  
* I don't suprise.

```{r}
small_verylarge <- diamonds %>%
  filter(carat < 1.0 | carat >= 2.5) %>%
  mutate(label = ifelse(carat < 1.0, "small", "verylarge"))

ggplot(data = small_verylarge, mapping = aes(x = price, y = ..density..)) +
  geom_freqpoly(mapping = aes(color = label))
```

4. Combine two of the techniques you’ve learned to visualise the combined distribution of cut, carat, and price.

* cut is a categorical variable, carat and price are continuous variable
  The following is one way that I can come up with, the trick is to use `x = cut_width(carat, 0.5)` rather
  than `x = carat`.

```{r}
smaller <- diamonds %>%
  filter(carat < 3)
ggplot(data = smaller, mapping = aes(x = cut_width(carat, 0.5), y = price, color = cut)) +
  geom_boxplot(mapping = aes(color = cut))
```

5. Two dimensional plots reveal outliers that are not visible in one dimensional plots. For example,
    some points in the plot below have an unusual combination of x and y values, which makes the points
    outliers even though their x and y values appear normal when examined separately.
    Why is a scatterplot a better display than a binned plot for this case?

* The outliers has crystallized in the following figure because it shows the relation between x and y.

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))
```

